---
title: "Data Preparation"
output: pdf_document
date: "2023-05-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Diffusion Model Analyses require certain response variables: subject number, response times in seconds, and whether the response was correct. 
As we are using existing data sets, some other variables are important as well, depending on the respective data set and manipulation (see short descriptions).

For usability reasons, variables will be renamed to follow one pattern:

- subject number - sub

- trial number - trial

- correct response - correct

- response time - RT

- further important variables will be renamed arbitrarily



# Chen 2019, Exp. 4a

## Short Description of Data / Experiment

Stimuli were presented in a matched form in study and test, and test lists were blocked by perceptual form (words and pictures were tested "separately"). Order of block was random and stimuli within a block were randomized. At test, participants were presented with an equal number of studied targets and unstudied foils randomly selected. Participants responded to the question "Have you studied this item?" by pressing a button, YES or NO, displayed below the test stimulus. All remaining details were identical to Experiments 1–3: Images and words were randomly intermixed during study.


## Clean Data

For our analyses with the diffusion model, we need the following variables:

- **subject** (SubNum)

- whether stimulus is a **word or an image** (StimType $\rightarrow$ stim)

- whether stimulus is **old or new** (StimIndex $\rightarrow$ item)

- participant's **response** (ResponseIndex $\rightarrow$ response)

- whether response is **correct** (Accurate)

- **response time** in seconds (RT)



```{r, include = FALSE}
# Load packages
library(tidyverse)

```


```{r}
# Load data
chen_4a <- read.csv2("chen-2019_Exp4a Words vs Images_SIR_Test.csv")

# Select and rename relevant variables
chen_4a_fin <- chen_4a %>% select("SubNum", "TestTrial", "StimType", "StimIndex", "ResponseIndex", "Accurate", "RT") %>% rename("sub" = "SubNum", "trial" = "TestTrial", "stim" = "StimType", "correct" = "Accurate") %>% mutate(StimIndex = ifelse(StimIndex == "targ", "old", "new")) %>% mutate(ResponseIndex = ifelse(ResponseIndex == "targ", "old", "new")) %>% rename ("response" = "ResponseIndex", "item" = "StimIndex")

# Show first rows
head(chen_4a_fin)

# Save new dataframe
write.csv(chen_4a_fin, "chen-2019-4a.csv")

```



# Chen 2019, Exp. 4b

## Short Description of Data / Experiment

See 4a. Stimuli were images and audio recordings instead of images and words. Images and audio recordings were randomly intermixed during study.


## Clean Data

For our analyses with the diffusion model, we need the following variables:

- **subject** (SubNum)

- whether stimulus is a **word or an image** (StimType)

- whether stimulus is **old or new** (StimIndex)

- participant's **response** (ResponseIndex)

- whether response is **correct** (Accurate)

- **response time** in seconds (RT)


```{r}
# Load data
chen_4b <- read.csv2("chen-2019_Exp4b Images vs Audio_SIR_Test.csv")

# Select and rename relevant variables
chen_4b_fin <- chen_4b %>% select("SubNum", "TestTrial", "StimType", "StimIndex", "ResponseIndex", "Accurate", "RT") %>% rename("sub" = "SubNum", "trial" = "TestTrial", "stim" = "StimType", "correct" = "Accurate") %>% mutate(StimIndex = ifelse(StimIndex == "targ", "old", "new")) %>% mutate(ResponseIndex = ifelse(ResponseIndex == "targ", "old", "new")) %>% rename ("response" = "ResponseIndex", "item" = "StimIndex")

# Show first rows
head(chen_4b_fin)

# Save new dataframe
write.csv(chen_4b_fin, "chen-2019-4b.csv")

```


# Criss 2017, Exp. 1

## Short Description of Data / Experiment

Each participant received two studytest lists, one mixed test and one blocked test, in random order. The study list was composed in a similar fashion for both conditions: 50 words and 50 faces were randomly selected from their respective pools and randomly ordered for each participant. Each item was studied for 3 s followed by a blank screen for 500 ms. The study was followed by a 45-s distractor task of simple addition. The test list contained 100 two alternative forced-choice (2AFC) test trials where participants were instructed to select the studied target from two choices, including the target and a randomly selected foil from the same stimulus type. The test was selfpaced and a 100-ms blank screen separated each trial.


## Clean Data

For our analyses with the diffusion model, we need the following variables:

- **subject** (subjectnum)

- whether stimulus is **old or new** (wordtype)

- whether response is **correct** (accuracy)

- **response time** in seconds (RT)

- whether stimulus is a **face or word** (faceorword; 0 = face, 1 = word,)

- condition - 1=blocked (words then faces), 2=blocked (faces then words), 3=mixed




```{r}
# Load data
criss <- read.csv2("criss-2017_expt1_osf.csv")

# Select and rename relevant variables
criss_fin <- criss %>% select("subjectnum", "testtrial", "condition", "faceorword", "accuracy", "RT") %>% rename("sub" = "subjectnum", "trial" = "testtrial", "correct" = "accuracy", "stim" = "faceorword")

# Show first rows
head(criss_fin)

# Save new dataframe
write.csv(criss_fin, "criss-2017.csv")

```




# Dedetas 2022

## Short Description of Data / Experiment

The experiment comprises two cycles. In each cycle, there were study and test phases which were separated with a distractor task. One cycle consisted of high-frequency words, and the other cycle consisted of low-frequency words. The order of word frequency lists was counterbalanced among participants: half of them received the high-frequency cycle first, followed by the low-frequency cycle while the other half received the low-frequency cycle first, followed by the high-frequency cycle.
In the study phase, participants were presented with 75 words, and in the test phase, they were tested on 150 words. The test list included 75 study words (targets) with 75 new words (foils). Participants indicated whether items were old or new. The word frequency was manipulated among cycles, meaning that in a cycle, there were only high or only low-frequency words presented in both study and test phases. There was no mixed condition since presenting high and low frequent words in random order in a test could interfere with the OI pattern of the test list.


## Clean Data

For our analyses with the diffusion model, we need the following variables:

- **subject** (SubID)

- whether stimulus is **old or new** (wordtype)

- participant's **response** (answertype)

- whether response is **correct** (correct)

- **response time** in seconds (RT)

- whether presented words are **high or low frequency words** (condition)


```{r}
# Load data
dedetas <- read.csv2("dedetas-2022_data.csv")

# Select and rename relevant variables
dedetas_fin <- dedetas %>% select("SubID", "condition", "testorder", "wordtype", "answertype", "correct", "RT") %>% rename("sub" = "SubID", "trial" = "testorder", "response" = "answertype", "item" = "wordtype")

# Show first rows
head(dedetas_fin)

# Save new dataframe
write.csv(dedetas_fin, "dedetas-2022.csv")

```











